%% Please fill in your name and collaboration statement here.
\newcommand{\studentName}{**FILL IN YOUR NAME HERE**}
\newcommand{\collaborationStatement}{**FILL IN YOUR COLLABORATION STATEMENT HERE \\ (See the syllabus for information)**}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[solution, letterpaper]{cscie121}
\usepackage{enumerate}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage{hyperref}
\usetikzlibrary{automata,positioning}
\begin{document}
\header{8}{November 20, 2015}



\problem{2+2+2+2+2+2}{2/3 page} 
For each part (A) through (F), answer TRUE or FALSE and briefly justify your answers.

\subproblem $n^6 - n^4 = \Theta(n^6 - 5n^3)$.

\subproblem $\sqrt{n} = \Omega(n^{\sin{n}})$.

\subproblem $n! = O(2^n)$.

\subproblem $n^{100001}=o(1.00001^n)$. (Hint: L'H\^opital's rule)

\subproblem $3^n=5^{O(n)}$.

\subproblem Suppose that $f(n) = g(O(n))$. Give a counterexample to prove that this does not necessarily imply that $f(n) = O(g(n))$, and show why your counterexample works.


\begin{solution}
  % Write your answer here.
\end{solution}


\problem{6} {1/3 page}
Two languages $H$ and $K$ are \emph{recursively separable} if there
exists some recursive language $R$ such that $H \subseteq R$ and
$K \subseteq \overline{R}$.  Prove that the languages $L_1 =
\{\langle M \rangle : M  \mbox{ accepts } \langle M \rangle\}$ and
$L_2 = \{\langle M \rangle: M \mbox{
rejects } \langle M \rangle \}$ are not recursively separable.

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{6}{1/3 page}
A function $f:\Sigma^* \rightarrow \Sigma^*$ is {\it computable} if
there exists a Turing Machine $M$ that, when given $s \in \Sigma^*$ as input,
halts with $f(s)$ written on the tape. The \emph{range} of $f$ is $\{f(x):
  x \in \Sigma^*\}$. Prove that a nonempty language is r.e.~if and only if 
  it is the range of a computable function. 

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{CHALLENGE! (3 extra credit points)}{1/2 page}
In this problem you will define a function that grows faster than any computable function. Thus you will prove the existence of an uncomputable function directly, without relying on Turing's diagonalization argument. 

The {\bf busy-beaver function}
$\beta(n)$ is
the largest number of $a$'s that can be printed by any $n$-state, two-symbol Turing machine that eventually halts when started from the empty tape.

Show that $\beta$ is not computable. 

\begin{solution}
  % Write your answer here.
\end{solution}


\problem{6+6}{1/2+2/3 pages}
Prove that the class ${\cal P}$ is closed under:
\subproblem Concatenation.
\subproblem Kleene star. (\emph{Hint:} Look at the algorithm we gave in class for recognizing context-free languages via Chomsky Normal Form.)

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{8}{.5 pages}
Prove that $\textsc{ALL}_\textsc{DFA} = \{\langle D \rangle : D$ is a DFA and $L(D)=\Sigma^*\}$ is in $\mathcal{P}$.

\begin{solution}
  % Write your answer here.
\end{solution}
\end{document}
