%% Please fill in your name and collaboration statement here.
\newcommand{\studentName}{**FILL IN YOUR NAME HERE**}
\newcommand{\collaborationStatement}{**FILL IN YOUR COLLABORATION STATEMENT HERE \\ (See the syllabus for information)**}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[solution, letterpaper]{cscie121}
\usepackage{enumerate}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage{hyperref}
\usetikzlibrary{automata,positioning}
\begin{document}
\header{7}{November 13, 2015}


\problem{4+4+4}{1 page}
Indicate whether each of the following languages is decidable. If it is
decidable, give a \textbf{high-level}\footnote{See Piazza for an explanation of
what a high-level description of a TM entails.} (but complete) description of
a Turing Machine that decides it. Otherwise, prove it is not decidable by
contradiction: show that if a decider $M_L$ existed for it, one could use $M_L$
to construct a new decider $M_L'$ that decides a problem we have proven to, in
fact, be undecidable (such as $\textsc{Halt}_{\textsf{TM}}$). You do not need
to define the mapping reduction function $f$ for this problem.\\

\noindent In all cases, let $T$ be a deterministic one-tape TM.
 
\subproblem $L=\{\left<T\right>:$ $T$ takes more than 53 steps on at least one
input.$\}$ 

\subproblem $L=\{\left<T\right>:$ $T$ takes more than 77 steps on at least one
input accepted by $T$.$\}$

\subproblem $L=\{\left<T\right>:$ $T$ takes more than 824 steps on all inputs
accepted by $T$.$\}$

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{3+3}{1/2 a page}
In the near future you're working as an engineer at Google/Microsoft/Facebook when your manager asks you to write the following two programs. Is this a problem? Provide a proof for your argument.

\subproblem Take another program's code as input and decide if that program is implemented with the shortest amount of code possible.

\subproblem Take another program's code and remove all inaccessible (dead) code from it.

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{4}{1/3 a page}

Let $A=\{\left<R,S\right>: \text{ $R$ and $S$ are regular expressions
and $L(R)
\subseteq L(S)$ \}}$. Show that $A$ is decidable by giving a high level description. 

\begin{solution}
  % Write your answer here.
\end{solution}



\problem{6}{1/3 page}

Prove that any decidable language $L$ is mapping reducible to any non-trivial decidable language $L'$. (Any decidable language other than $\Sigma^*$, $\emptyset$)

\begin{solution}
  % Write your answer here.
\end{solution}


\problem{6}{1/3 page}

For any Turing Machine $M$, the language accepted by $M$, $L(M)$ is always
recursively enumerable. Let $M_F$ be the machine created by flipping the accept
and reject states of $M$. Prove that $L(M_F)$ is not always recursive whenever $L(M)$ is recursive.

\begin{solution}
  % Write your answer here.
\end{solution}


\problem{6}{1/3 page}
A computation of a Turing Machine \emph{loops} if it repeats a configuration, that
is, re-enters the same state with the identical tape and the head
position. Prove that if every non-halting computation of $M$ loops, then $L(M)$ is decidable.
In particular, you need to give a high level description of a Turing machine that 
decides $L(M)$, and provide a short justification. 

\begin{solution}
  % Write your answer here.
\end{solution}




\end{document}
