%% Please fill in your name and collaboration statement here.
\newcommand{\studentName}{**FILL IN YOUR NAME HERE**}
\newcommand{\collaborationStatement}{**FILL IN YOUR COLLABORATION STATEMENT HERE \\ (See the syllabus for information)**}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[solution, letterpaper]{cscie121}
\usepackage{enumerate}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage{hyperref}
\usetikzlibrary{automata,positioning}
\begin{document}

\header{6}{November 6, 2015}


\problem{6}{1/3 page}
Prove that the class of decidable languages is closed under union. 

\begin{solution}
  % Write your answer here.
\end{solution}


\problem{6}{1/3 page}
Prove that the class of Turing-recognizable languages is closed under intersection.

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{3 + 5}{5 lines, 1/3 page}
\subproblem Consider $L = \{\langle M \rangle: M \text{ is a TM that accepts no strings shorter than } 42 \text{ characters in length}\}$. 
 
Prove that L is not decidable. Hint: Consider Rice's Theorem.

\subproblem Use that the fact that $L$ is not decidable to prove that $L$ is not recognizable.

\begin{solution}
  % Write your answer here.
\end{solution}


\problem{3}{5 lines}
Define $\mathsc{Prefix}(L) = \{ x : xy \in L \text{ for some }y \in \Sigma^*\}$. Show that if $L$ is Turing-recognizable, then $\mathsc{Prefix}(L)$ is Turing-recognizable.

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{8} {2/3 page}
Let $L=\{\langle M\rangle:$ $M$ moves left on the tape at some point when run on $\varepsilon\}$. Show that $L$ is decidable.

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{10} {3/4 page}
Show that a language $L_1$ is Turing-recognizable if and only if there exists some decidable language $L_2$ such that $L_1 = \{x : \text{there exists $y$ such that $\langle x,y \rangle \in L_2$}\}$. \\ \\
Note that the $\langle \rangle$ notation signifies that you are ``stringify-ing'' the contents within the bracket, since Turing Machines take strings as input. \\ \\ 
(Hint: Imagine that $y$ gives you information about an accepting computation on $x$, if one exists.) \\ \\
\begin{solution}
  % Write your answer here.
\end{solution}

\problem{Challenge!! (3)}{1/3 page}
Show that every infinite regular language has a
subset that is Turing-recognizable but undecidable.

\begin{solution}
  % Write your answer here.
\end{solution}

\end{document}

