%% Please fill in your name and collaboration statement here.
\newcommand{\studentName}{**FILL IN YOUR NAME HERE**}
\newcommand{\collaborationStatement}{**FILL IN YOUR COLLABORATION STATEMENT HERE \\ (See the syllabus for information)**}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[solution, letterpaper]{cscie121}
\usepackage{enumerate}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage{hyperref}
\usetikzlibrary{automata,positioning}
\begin{document}
\header{5}{Due October 30, 2015 11:59pm}



\problem{4+2}{1/2 page} Consider the Turing machine $M=(Q, \Sigma,
\Gamma,\delta, q_0, q_{\mathrm{accept}}, q_{\mathrm{reject}})$
where\begin{itemize}\item
$Q = \{q_0, q_1, q_2, q_{\mathrm{accept}},q_{\mathrm{reject}}\}$,
\item $\Sigma = \{a,b\}$ and $\Gamma =\{a,b,\sqcup\}$, \item The
start, accept and reject states are $q_0$, $q_{\mathrm{accept}}$,
and $q_{\mathrm{reject}}$ respectively. \item The function
$\delta$ is given
by:\end{itemize}\begin{center}\begin{tabular}{cc|r}$q$ & $\sigma$
& $\delta(q, \sigma)$ \\ \hline\hline$q_0$ & $a$ & $(q_0, b,
R)$\\$q_0$ & $b$ & $(q_1, a, R)$\\$q_0$ & $\sqcup$ &
$(q_{\mathrm{reject}}, \sqcup, R)$\\$q_1$ & $a$ & $(q_2, b,
R)$\\$q_1$ & $b$ & $(q_1, a, R)$\\$q_1$ & $\sqcup$ &
$(q_{\mathrm{reject}}, \sqcup, R)$\\$q_2$ & $a$ & $(q_2, b,
R)$\\$q_2$ & $b$ & $(q_2, a, R)$\\$q_2$ & $\sqcup$ &
$(q_{\mathrm{accept}}, \sqcup,
R)$\\\end{tabular}\end{center}

\subproblem Give the sequence of configurations describing $M$'s computation
on the string $aabba$

\subproblem Describe $L(M)$.


\begin{solution}
  % Write your answer here.
\end{solution}

\pagebreak

\problem{4}{1/4 page}
A Modern Turing Machine (MTM), instead of $\{L,R\}$, has $\{L_k, R_k\}$ (move the head left or right $k$ spaces on the tape, for any integer $k$ encoded in the rule in the MTM). Prove that the MTMs are equivalent in power to the TMs. In other words, any MTM can be converted to a TM with an equivalent language, and vice versa.

\begin{solution}
  % Write your answer here.
\end{solution}


\problem{2 + 2 + 1}{1/2 page}

A Democratic rule of a CFG $(V, \Sigma, R, S)$ is one whose right-hand side is a member of $V\Sigma^*\cup\Sigma^*$, that is, only the leftmost symbol can be a nonterminal. Republican rules are defined analogously where only the rightmost symbol can be a nonterminal. Show that a grammar with only Democratic rules, or only Republican rules, generates a regular language, but a grammar with a mixture of Democratic and Republican rules may generate a non-regular language.

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{7}{3/4 page} Let us define EvenPalindrome$(L)=\{ww^{R} : w \in
L\}$. Show that if $L$ is regular, then EvenPalindome$(L)$ is
context-free.  Note that this is not the same as the language of all even palindromes, which is actually just EvenPalindrome($\Sigma^*$). Explain briefly how your solution works. You need not provide a formal proof of correctness.

\begin{solution}
  % Write your answer here.
\end{solution}


\problem{CHALLENGE!! (3)} {1/2 page}
A Boring Turing Machine (BTM) can only write $\#$ on the tape (assume $\# \notin \Sigma$). 
Prove that the BTMs are equivalent in power to the TMs.

\begin{solution}
  % Write your answer here.
\end{solution}

\problem{6} {1/3 page}

\noindent Show that every infinite Turing-recognizable language
has an infinite decidable sublanguage.

\begin{solution}
  % Write your answer here.
\end{solution}



\problem{4+2+4}{1 page}

Let $G = (V, \Sigma, R, S)$ where $V=\{S,V\}$, $\Sigma = \{a,b\}$, and $R$ is the set of rules: 

$$ 
S \rightarrow bSS \| aS \| aV $$
$$ V \rightarrow aVb \| bVa \| VV \| \varepsilon 
$$

\subproblem
Transform $G$ into an equivalent grammar $G'$ in Chomsky normal form. Show your work for each step of this conversion, but long justifications for each step are not necessary.

\subproblem 
Explain in English what language the grammar $G$ generates. (One clearly worded sentence is fine; thorough justification is not necessary, but it might get you some partial credit if your answer is wrong.)

\subproblem
Check whether the strings $abaab$ and $bbabaa$ are generated by $G$, using the recognition algorithm for grammars in Chomsky normal form given in class. Show the complete filled-in matrix for each of the two strings.


\begin{solution}
  % Write your answer here.
\end{solution}
\end{document}
